---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dcatapchr


The goal of dcatapchr is to create a metadata catalog compatible with the 
DCAT-AP CH standard.

See for example https://www.dcat-ap.ch/releases/2.0/dcat-ap-ch.html and https://handbook.opendata.swiss/de/content/glossar/bibliothek/dcat-ap-ch.html  

See https://ckan.opendata.swiss/api/3/action/harvest_source_list for catalog examples. See https://www.web.statistik.zh.ch/ogd/harvester/stadt-winterthur.xml or https://www.web.statistik.zh.ch/ogd/harvester/statistisches-amt-kanton-zuerich.xml  




## Classes of DCAT-AP CH

- Catalog, `dcat:Catalog`, mandatory  
- Dataset, `dcat:Dataset`, mandatory  
- Distribution, `dcat:Distribution`, mandatory  

Properties of `dcat:Catalog`:  
- dataset, `dcat:dataset`, mandatory  


Properties of `dcat:Dataset`:  
- title, `dct:title`, mandatory, multilingual  
- description, `dct:description`, mandatory, multilingual  
- publisher, `dct:publisher`, mandatory, import format changed  
- contact point, `dcat:contactPoint`, mandatory  
- identifier, `dct:identifier`, mandatory  
- distribution, `dcat:distribution`, mandatory  
- ... (other properties, conditional or optional)  



Properties of `dcat:Distribution`:  
- issued, `dct:issued`, mandatory  
- access url, `dcat:accessURL`, mandatory  
- rights, `dct:rights`, mandatory  
- ... (other properties, conditional or optional)  


The properties above are those of DCAT-AP CH V1 (I think).  


## DCAT-AP CH V2

There is now V2:  
- https://www.dcat-ap.ch/  
- https://www.dcat-ap.ch/releases/2.0/dcat-ap-ch.html  

- https://www.dcat-ap.ch/releases/2.0/dcat-ap-ch.html#changes-overview  
- https://www.dcat-ap.ch/releases/2.0/dcat-ap-ch.html#dcat-ap-building-blocks  

- The Requirement level "conditional" was replaced by the "recommended", to align with the DCAT-AP's notation  

**Classes**:  
- Catalog --> `dcat:Catalog`  
- Dataset --> `dcat:Dataset`  
- Distribution --> `dcat:Distribution`  
- Data Service --> `dcat:DataService`  

See https://www.dcat-ap.ch/releases/2.0/dcat-ap-ch.html#classes  

### Mandatory (and some recommended) properties (V2)

Properties of `dcat:Catalog`:  
- description: `dct:description` (mandatory)  
- publisher: `dct:publisher` (mandatory)  
- title: `dct:title` (mandatory)  
- dataset, `dcat:dataset`, recommended (V1: mandatory)  
- ... (other properties, recommended)  
- homepage, `foaf:homepage`, recommended  
- language, `dct:language`, recommended  
- issued, `dct:issued`, recommended  
- rights, `dct:rights`, recommended  
- service, `dcat:service`, recommended  
- spatial, `dct:spatial`, recommended  
- themeTaxonomy, `dcat:themeTaxonomy`, recommended  
- modified, `dct:modified`, recommended  




Properties of `dcat:Dataset`:  
- title, `dct:title`, mandatory, multilingual  
- description, `dct:description`, mandatory, multilingual  
- publisher, `dct:publisher`, mandatory, import format changed  
- contact point, `dcat:contactPoint`, mandatory  
- identifier, `dct:identifier`, mandatory  
- distribution, `dcat:distribution`, recommended (V1: mandatory)  
- ... (other properties, recommended)  
- keyword, `dcat:keyword`, recommended  
- landingPage, `dcat:landingPage`, recommended  
- issued, `dct:issued`, recommended  
- spatial, `dct:spatial`, recommended  
- temporal, `dct:temporal`, recommended  
- theme, `dcat:theme`, recommended  
- modified, `dct:modified`, recommended  



Properties of `dcat:Distribution`:  
- access url, `dcat:accessURL`, mandatory  
- license, `dct:license`, mandatory  
- rights, `dct:rights`, recommended (V1: mandatory)  
- ... (other properties, recommended)  
- availability, `dcatap:availability`, recommended  
- description, `dct:description`, recommended  
- format, `dct:format`, recommended  
- title, `dct:title`, recommended  
- modified, `dct:modified`, recommended  



## Installation

You can install the development version of dcatapchr like so:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(dcatapchr)

# Create the metadata of a dataset with 3 distributions

# First create the 3 distributions
distributions <- c(
  dcat_distribution(
    issued = dct_issued("2021-10-14"), 
    access_url = dcat_accessURL(
      "https://www3.ti.ch/DFE/DR/USTAT/allegati/cubo/cubi_POL_01_xlsx.zip"
    ), 
    rights = dct_rights(
      "NonCommercialAllowed-CommercialAllowed-ReferenceRequired"
    )
  ),
  
  dcat_distribution(
    issued = dct_issued("2021-10-14"), 
    access_url = dcat_accessURL(
      "https://www3.ti.ch/DFE/DR/USTAT/allegati/cubo/cubi_POL_01_px.zip"
    ), 
    rights = dct_rights(
      "NonCommercialAllowed-CommercialAllowed-ReferenceRequired"
    )
  ),
  
  dcat_distribution(
    issued = dct_issued("2021-10-14"), 
    access_url = dcat_accessURL(
      "https://www3.ti.ch/DFE/DR/USTAT/allegati/cubo/cubi_POL_01_csv.zip"
    ), 
    rights = dct_rights(
      "NonCommercialAllowed-CommercialAllowed-ReferenceRequired"
    )
  )
)

distributions

# Then create the dataset with the 3 distributions created above
description <- readLines(
  system.file(
    "cubi_POL_01_INFORMAZIONI.txt", package = "dcatapchr"
  ), encoding = "UTF-8"
)
description <- strwrap(description, width = 80)
description <- paste(description, collapse = "\n")

# Substitute & with &amp; in urls
description <- gsub("&", "&amp;", description, fixed = TRUE)

dataset <- dcat_dataset(
  title = dct_title(
    title = "Partecipazione alle elezioni del Consiglio di Stato, per comune, in Ticino, dal 1921 al 2011", 
    language = "it"
  ), 
  description = dct_description(
    description = description, 
    language = "it"
  ), 
  publisher = dct_publisher(publisher = "https://www.ti.ch/ustat", foaf_name = "Ustat"), 
  contactPoint = dcat_contactPoint(), 
  identifier = dct_identifier(),
  distribution = distributions
)

# dataset

# And finally put the dataset in a catalog
catalog <- dcat_catalog(
  catalog_endpoint = "https://www3.ti.ch/DFE/DR/USTAT/allegati/digitale/catalogo_ustat.rdf", 
  datasets = dataset
)

head(catalog, n = 15)


tail(catalog, n = 20)
```

Export to an rdf file:
```{r eval=FALSE}
f <- file("./catalogo_ustat.rdf", "wb")
writeLines(catalog, con = f, useBytes = TRUE)
close(f)
```

